Palomar 5 Gaps

### 2024-09-19 paper/main.tex
added change in momentum

  The goal of this study is to contribute to the ongoing effort to use stellar streams as probes for dark matter on sub-galactic scales. According to the cold dark matter model, which is supported by observations, galaxies form hierarchically, leading to a mass distribution of dark matter subhalos that follows a power-law. This suggests that numerous small dark matter clumps should exist within galaxies. \citet{rodrigo_ibata_uncovering_2002} suggested that the presence of said dark matter subhaloes dark matter subhalos can influence the stellar streams by diffussing their distribution of orbital elements. \citet{r_g_carlberg_pal_2012} extended this idea by noting that subhaloes could have flybys and that leave gaps in stellar streams and subsquently explored the statistical distribution the number of gaps gaps given a probabilistic model of the rate of encounters given based on the expected distribution of dark matter subhalos. \citet{bonaca_spur_2018} took a more direct observational approach, identifying an underdensity in the GD-1 stream that could not be explained by known objects like globular clusters and was inconsistent with the mass-size relationship predicted by the Lambda-CDM model.

@@

    The primary methodological departure from Ferrone et al. (2023) is that now we include globular cluster interactions. Instead of treating the clusters as point masses that only experience the gravitational potential of the galaxy, we now compute N-body interactions between the clusters during the backward integration over 5 billion years. Although these interactions were included, we observed no significant exchanges of orbital energy or momentum between clusters. The resulting orbits were saved to data files for subsequent analysis.

    At the computed positions 5 Gyr ago, each globular cluster was replaced by a system of 100,000 particles, sampled from a Plummer distribution to represent the stellar content of the cluster. We then integrated the evolution of these particles forward in time to the present day. During this forward integration, each star particle feels the gravitational influence of the galaxy and all globular clusters, but not other star particles. We do not perform direct N-body simulations between star particles; instead, at each time step, we compute the positions of all globular clusters and apply their forces to the star particles.

@@

    \begin{equation}
        \begin{aligned}
          \Delta \vec{v} &=\vec{v_p} - \vec{v_s} \\
          w_\parallel &= \left(\Delta \vec{v}\right)\cdot \hat{v_s}\\  
          w_\perp &=  \sqrt{\Delta v ^2 - w_\parallel ^ 2}
          \end{aligned}
        \end{equation}

@@

  \begin{figure}
    \centering
    \includegraphics[width=1\linewidth]{E_Lz_perturbers.png}
    \includegraphics[width=1\linewidth]{mass_size_plane.png}
    \caption{One plot shows the distribution in energy and angular momentum, the next are the inernal proeprties of the clusters, and the last show when the clusters commit their crime, and how many times it happens accross the 50 simulations. }
    \label{fig:mass_size_plane}
    \end{figure}
  \begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{histogram_impact_time.png}
    \caption{When the impacts occured}
    \label{fig:histogram_impact_time}
    \end{figure*}

@@

>>>>

  The goal of this study is to contribute to the ongoing effort to use stellar streams as probes for dark matter on sub-galactic scales. According to the cold dark matter model, which is supported by observations, galaxies form hierarchically. As shown in Springel et al 2008 in the suite of Acquarious simulations, $\lambda$-CDM predicts a distribution of masses of dark matter subhalos that follows a power-law. The authors also demonstrate that subhaloes should exist within other haloes. To my knowledge, thanks to the introduction of Bonaca et al 2019 who pointed this paper out to me, the smallest dark matter sub halo detected to date was found within an enstien ring from gravitational lensing by Vigetti et al 20XX.

  Over the years, people began to think of the role that subhalos can have on effecting stellar streams. \citet{rodrigo_ibata_uncovering_2002} suggested that the presence of said dark matter subhaloes dark matter subhalos can influence the stellar streams by diffussing their distribution of orbital elements. \citet{r_g_carlberg_pal_2012} extended this idea by noting that subhaloes could have flybys and that leave gaps in stellar streams and subsquently explored the statistical distribution the number of gaps gaps given a probabilistic model of the rate of encounters given based on the expected distribution of dark matter subhalos. \citet{bonaca_spur_2018} took a more direct observational approach, identifying an underdensity in the GD-1 stream that could not be explained by known objects like globular clusters and was inconsistent with the mass-size relationship predicted by the Lambda-CDM model.

@@

    The primary methodological departure from Ferrone et al. (2023) is that now we include globular cluster interactions. Instead of treating the clusters as point masses that only experience the gravitational potential of the galaxy, we now compute N-body interactions between the clusters during the backward integration over 5 billion years. The equation of motion for the globular clusters is thus: 
    \begin{equation}
      \ddot{\vec{r}}_i = -\nabla \Phi(R_i,z_i) + \left.\sum_{j\neq i} \frac{Gm_j}{\left(|\vec{r}_j - \vec{r}_i|^2 + b_j^2\right)^{3/2}}\right. \left(\vec{r}_j - \vec{r}_i\right).
      \end{equation}
    Although these interactions were included, we observed no significant exchanges of orbital energy or momentum between clusters. The resulting orbits were saved to data files for subsequent analysis.

    At the computed positions 5 Gyr ago, each globular cluster was replaced by a system of 100,000 particles, sampled from a Plummer distribution to represent the stellar content of the cluster. We then integrated the evolution of these particles forward in time to the present day. During this forward integration, each star particle feels the gravitational influence of the galaxy and all globular clusters, but not other star particles. We do not perform direct N-body simulations between star particles; instead, at each time step, we compute the positions of all globular clusters and apply their forces to the star particles. The equation of motion is thus: 
    \begin{equation}
      \ddot{\vec{r}}_p = -\nabla \Phi(R_p,z_p) + \left.\sum_{j} \frac{Gm_j}{\left(|\vec{r}_j(t) - \vec{r}_p|^2 + b_j^2\right)^{3/2}}\right. \left(\vec{r}_j(t)- \vec{r}_p\right)
      \end{equation}

@@

    \begin{equation}
        \begin{aligned}
          \delta \vec{v} &=\vec{v_p} - \vec{v_s} \\
          w_\parallel &= \left(\delta \vec{v}\right)\cdot \hat{v_s}\\  
          w_\perp &=  \sqrt{\Delta v ^2 - w_\parallel ^ 2}
          \end{aligned}
        \end{equation}

@@

  \begin{figure}
    \centering
    \includegraphics[width=1\linewidth]{E_Lz_perturbers.png}
    \includegraphics[width=1\linewidth]{mass_size_plane.png}
    \includegraphics[width=1\linewidth]{impact_geometry_statistics_deltaP.png}
    \caption{One plot shows the distribution in energy and angular momentum, the next are the inernal proeprties of the clusters, and the last show when the clusters commit their crime, and how many times it happens accross the 50 simulations. }
    \label{fig:mass_size_plane}
    \end{figure}
  \begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{histogram_impact_time.png}
    \caption{When the impacts occured}
    \label{fig:histogram_impact_time}
    \end{figure*}

@@

  \section{Impact Geometry Statistics}

  \begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{impact_geometry_statistics.png}
    \caption{Impact geometry statistics}
    \label{fig:impact_geometry_statistics}    
    \end{figure*}
